On centrality of orthogonal $\Kt$

### 2016-04-29 paper.tex
Minor tweaks in intro

\title{On centrality of even orthogonal $\Kt$}
\keywords {Steinberg groups, $K_2$-functor. {\em Mathematical Subject Classification (2010):} 19C09}

@@

Recall that for each reduced root system $\Phi$ and a commutative ring $R$ one can associate a split simple group $\GG(\Phi, R)$ called the \emph{Chevalley group}.
For $\Phi$ of rank $\geq 2$ the abstract group $\GG(\Phi, R)$ contains a large normal subgroup $\E(\Phi, R)$ generated by the elementary root unipotents $t_\alpha(\xi)$, $\xi\in R$, $\alpha\in \Phi$.
The quotient group functor $\Ko(\Phi, R)=\GG(\Phi, R)/\E(\Phi, R)$ is called the \emph{unstable $\Ko$-functor}.
This functor shares many properties with its stable counterpart $\mathrm{SK}_1(R) = \lim\limits_{l\to\infty}\Ko(\rA_\ell, R)$ which, in turn, is an essential direct summand of $\Ko(R)$.

@@

The special case $\Phi=\rA_\ell$, $\ell\geq 4$ of \cref{tm:lg-principle} was demonstrated for the first time by Tulenbaev, see~\cite[Theorem~2.1]{Tul},
while similar assertions for $\Phi=\rC_\ell$, $\ell\geq 3$ and $\Phi=\rE_\ell$, $\ell=6,7,8$ have been established in \cite{Lav2} and \cite{SCh}, respectively.
The special cases $\Phi=\rA_3$ and $\Phi=\rD_\ell$, $\ell\geq 4$ of \cref{tm:lg-principle} are new and have not been known before.

There are known counterexamples to centrality of $\Kt$ in the rank $2$ case (see~\cite{W}).
This shows that the assumption on the rank of $\Phi$ in \cref{tm:lg-principle} is strict.

@@

Our proof of \cref{tm:lg-principle} follows the same lines as in \cite[Theorem~1]{SCh}.
More concretely, using the same patching technique as in \cite{SCh} we reduce the problem to the case $\Phi=\rA_3$.
Then, we formulate and prove a new presentation of the linear group $\St(4, R)$ and then utilize it to reprove 
Tulenbaev's key lemma (cf. \cite[Lemma~2.3]{Tul}) in the rank $3$ case.

The rest of the paper is organised as follows. 
In \cref{sec:yap} we formulate ``yet another presentation'' for the rank $3$ Steinberg groups.
%It mimics the standart vector representation of $\SL(n,\,R)$, therefore in the first section we use representations and corresponding vector notation. 
Finally, in \cref{sec:patching} we prove our main results.
%There we switch to the root notation.

@@

\section{Patching Tulenbaev maps}\label{sec:patching}

The next section is devoted to the proof of Theorem~\ref{a3map} for Steinberg groups corresponding to an arbitrary simply-laced root system of rank $l\geq3$. To do so, we will ``glue'' maps $T$ constructed in the present section for systems of type $\mathrm A_3$. All main results of the present paper follow from the existence of such a map.

>>>>

\title{On centrality of orthogonal $\Kt$}
\keywords {Steinberg groups, $K_2$-functor. {\em Mathematical Subject Classification (2010):} 19C09}

@@

Recall that to each reduced root system $\Phi$ and a commutative ring $R$ one can associate a split simple group $\GG(\Phi, R)$ called the \emph{Chevalley group}.
For $\Phi$ of rank $\geq 2$ the abstract group $\GG(\Phi, R)$ contains a large normal subgroup $\E(\Phi, R)$ generated by the elementary root unipotents $t_\alpha(\xi)$, $\xi\in R$, $\alpha\in \Phi$.
The quotient group $\Ko(\Phi, R)=\GG(\Phi, R)/\E(\Phi, R)$ is called the \emph{unstable $\Ko$-group}.
This group shares many properties with its stable counterpart $\mathrm{SK}_1(R) = \lim\limits_{l\to\infty}\Ko(\rA_\ell, R)$ which, in turn, is an essential direct summand of $\Ko(R)$.

@@

The special case $\Phi=\rA_\ell$, $\ell\geq 4$ of \cref{tm:lg-principle} was demonstrated for the first time by Tulenbaev, see~\cite[Theorem~2.1]{Tul},
while cases $\Phi=\rC_\ell$, $\ell\geq 3$ and $\Phi=\rE_\ell$, $\ell=6,7,8$ have been analyzed in \cite{Lav2} and \cite{SCh}, respectively.
The special cases $\Phi=\rA_3$ and $\Phi=\rD_\ell$, $\ell\geq 4$ of \cref{tm:lg-principle} are new and have not been known before.

There are known counterexamples to centrality of $\Kt$ in rank $2$ (see~\cite{W}).
This shows that our assumption on the rank of $\Phi$ in \cref{tm:lg-principle} is strict.

@@

Our proof of \cref{tm:lg-principle} is similar to that of \cite[Theorem~1]{SCh}.
Using the same patching technique we reduce the problem to the case $\Phi=\rA_3$.
Then, using a new presentation of the relative Steinberg group $\St(4, R, I)$ we reprove Tulenbaev's key lemma (cf. \cite[Lemma~2.3]{Tul}) in the rank $3$ case.

The rest of the paper is organised as follows. 
In \cref{sec:yap} we describe a new symmetric presentation for the relative Steinberg group of rank $3$.
This presentation is then used in~\cref{sec:lgp} to establish the particular case $\Phi=\rA_3$ of \cref{tm:lg-principle}.
%It mimics the standart vector representation of $\SL(n,\,R)$, therefore in the first section we use representations and corresponding vector notation. 
Finally, in \cref{sec:patching} we derive our principal results from this special case and the amalgamation theorem for relative Steinberg groups from~\cite{SCh}.
%There we switch to the root notation.

@@

\section{Patching Tulenbaev maps (to be written)}\label{sec:patching}

The next section is devoted to the proof of Theorems~\ref{tm:centrality}--\ref{tm:lg-principle} for Steinberg groups corresponding to 
arbitrary simply-laced root system of rank $l\geq3$.
To do so, we glue maps $T$ constructed in the previous section following the approach of~\cite{SCh}.
